---
title: "Adding COinS Metadata for Zotero"
description: "Understand COinS metadata and how add_coins() helps reference managers like Zotero automatically import your blog posts."
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Adding COinS Metadata for Zotero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is COinS?

**COinS** stands for **ContextObjects in Spans**. It's a simple, standardized way to embed bibliographic metadata directly into HTML pages so that specialized software can read and extract it.

Think of it like this: when you visit a website, humans see the text and can understand what a blog post is about. But reference managers like Zotero need machine-readable metadata to know the title, author, publication date, and other citation details without requiring manual data entry.

COinS solves this by embedding bibliographic information in an invisible HTML `<span>` element on your page. When Zotero visits your blog post, it detects the COinS metadata and can automatically add the post to your library with all the correct fields filled in.

---

## Why Use COinS?

### The Problem

Without COinS, Zotero has to guess your post's metadata by examining the HTML structure. This often fails for blog posts because:

- Blog post titles might be in an `<h1>` tag, but so might many other things
- The publication date might be hidden in an obscure `<time>` element or embedded in a longer text string
- Authors are rarely marked up consistently across different blog platforms
- Categories or tags have no standard HTML representation

### The Solution

COinS eliminates guessing. When you add a COinS span to your post, Zotero can:

1. **Automatically detect the post's metadata** — no manual entry needed
2. **Save the post to your personal library** — with correct author, date, title, and keywords
3. **Use correct citation formatting** — because Zotero knows the post is a blog article, not a book or journal article
4. **Reference it in your writing** — Zotero generates correct citations in your preferred style (Chicago, APA, etc.)

This is especially valuable for academic writers who rely on reference managers for literature management.

---

## The Metadata in COinS

A COinS entry is built from key-value pairs that describe a published work. The main fields for a blog post are:

| Field | Example | Purpose |
|-------|---------|---------|
| `rft.title` | "My Amazing Discovery" | Article title |
| `rft.au` | "Peter Baumgartner" | Author name |
| `rft.date` | "2026-08-22" | Publication date |
| `rft.genre` | "blogPost" | Content type (in this case, a blog post) |
| `rft_val_fmt` | "info:ofi/fmt:kev:mtx:journal" | Standard identifier for the metadata format |
| `rft.description` | "A brief summary..." | Post description or abstract |
| `rft.keywords` | "research, data science" | Categories or tags |
| `ctx_ver` | "Z39.88-2004" | COinS specification version |

Additional optional fields can include:
- `rft.pub` — Publication/blog name
- `rft.spage`, `rft.epage` — Page numbers (not applicable to blog posts)
- `rft.issn` — Journal ISSN (not applicable to blog posts)
- `rft.language` — Language code (e.g., "en-US")

---

## The Standard Behind COinS

COinS is based on the **OpenURL** standard (ANSI/NISO Z39.88-2004), which was originally designed to help libraries link users to full-text content. The "ContextObject" is the standardized metadata package at the heart of OpenURL.

The clever part of COinS is that it **embeds an OpenURL ContextObject invisibly in HTML** using a `<span>` element:

```html
<span class="Z3988" title="ctx_ver=Z39.88-2004&rft.title=...&rft.au=..."></span>
```

The `class="Z3988"` tells processors like Zotero that this span contains COinS metadata. The actual data is in the `title` attribute (which is counterintuitive—it's not displayed as a tooltip; it's just a container for the encoded metadata).

### Why a span?

Using a `<span>` allows COinS metadata to be placed anywhere in a web page without breaking the HTML structure or being visible to human readers. Other approaches (like storing metadata in `<meta>` tags in the page's `<head>`) are more restrictive—you can only have one set of metadata per page. With COinS, you can embed metadata for multiple items on the same page.

---

## Websites Using COinS

COinS has been adopted by many major platforms and content providers:

- **Wikipedia** — embedded in citation templates and bibliographic data
- **Academic repositories** — CiteULike, Citebase, Hubmed, ResearchGate
- **Library catalogs** — Copac, WorldCat, VuFind
- **Publishing platforms** — WordPress blogs, Institutional repositories
- **Specialized services** — Zotero itself can generate COinS for shared bibliographies

---

## How Zotero Uses COinS

When you're browsing a blog post with COinS in your Zotero-enabled browser:

1. Zotero's browser extension scans the page and finds the `<span class="Z3988">` element
2. It decodes the embedded metadata
3. It displays a **save icon** (📖) in your browser address bar, indicating that Zotero can capture the page
4. You click the icon to save the post to your library
5. Zotero populates the entry with all the metadata from COinS: title, author, date, description, etc.

Without COinS, Zotero might not recognize the page as a citable item, and you'd have to manually enter all the metadata.

---

## Adding COinS to Your Blog Posts

The `add_coins()` function automates COinS generation. Here's how to use it:

### Prerequisites

- A completed Quarto blog post in `posts/your-post/index.qmd`
- A `_quarto.yml` file in your blog's root directory with at least `site-url:` specified
- Optional: configuration in your project's `.Rprofile`

### Basic Usage

```r
library(qpost)

# Add COinS metadata to the currently open post in RStudio
add_coins()

# Or specify a file path explicitly
add_coins(file_path = "posts/my-post/index.qmd")
```

### With configuration

Set up your blog's metadata in the project `.Rprofile`:

```r
# Project-level .Rprofile
options(
  qpost.lang = "en",          # Language code (auto-detected from _quarto.yml if present)
  qpost.license = "CC-BY-4.0" # Creative Commons license for your blog posts
)
```

Then call `add_coins()` as above.

---

## What add_coins() Does

When you call `add_coins(file_path)`:

1. **Reads your post's YAML front matter** — extracts title, author, date, categories, description, etc.
2. **Queries `_quarto.yml`** — reads the blog's site URL and language settings
3. **Applies auto-resolution** — fills in missing fields from defaults:
   - `language` — from post YAML → from `_quarto.yml` → from `getOption("qpost.lang")` → omitted
   - `license` — from post YAML → from `getOption("qpost.license")` → omitted
4. **Generates the COinS metadata** — creates the properly formatted key-value string
5. **Creates an R code chunk** — appends a chunk to your post that displays the metadata
6. **Opens the edited post** — so you can review and commit the changes

### Generated Output

At the bottom of your `index.qmd`, `add_coins()` appends:

```{r}
#| label: coins
#| echo: false
#| results: asis

cat('<span class="Z3988" title="ctx_ver=Z39.88-2004&rft_val_fmt=info%3Aofi%2Ffmt%3Akev%3Amtx%3Ajournal&rft.genre=blogPost&rft.atitle=My%20Amazing%20Discovery&rft.au=Peter%20Baumgartner&rft.date=2026-08-22&rft.description=A%20brief%20summary...&rft.language=en&rft.keywords=research%2Cdata%20science&rft.license=CC-BY-4.0"></span>')
```

This code:
- Uses an R chunk with `echo: false` and `results: asis` so it's invisible to readers
- Outputs the COinS `<span>` as raw HTML
- Is placed **at the end** of your post, so Zotero sees the complete content first

---

## The Metadata in Detail

Here's what each key means in the generated COinS:

```
ctx_ver=Z39.88-2004
```
Specifies the COinS/OpenURL version.

```
rft_val_fmt=info:ofi/fmt:kev:mtx:journal
```
Declares the metadata format as "key-encoded-value" (`kev`) for a journal-like resource. We use this for blog posts because they're closest to journal articles in structure.

```
rft.genre=blogPost
```
Identifies the resource type as a blog post (not a book, journal article, or conference paper).

```
rft.atitle=My%20Amazing%20Discovery
```
The article title (URL-encoded, so spaces become `%20`).

```
rft.au=Peter%20Baumgartner
```
The author name.

```
rft.date=2026-08-22
```
Publication date in ISO format (`YYYY-MM-DD`).

```
rft.description=A%20brief%20summary...
```
Post description or abstract (optional).

```
rft.language=en
```
Language code (optional). Defaults to your blog's language from `_quarto.yml`.

```
rft.keywords=research%2Cdata%20science
```
Categories/tags from your post (comma-separated, URL-encoded).

```
rft.license=CC-BY-4.0
```
License identifier (optional). Set in `.Rprofile` to apply the same license to all posts.

---

## Configuration

### Auto-resolution Hierarchy

`add_coins()` builds metadata using this priority order:

| Field | Source priority |
|-------|-----------------|
| **title** | Post YAML `title:` (required) |
| **date** | Post YAML `date:` (required) |
| **author** | Post YAML `author:` |
| **description** | Post YAML `description:` |
| **categories/keywords** | Post YAML `categories:` |
| **language** | (1) Post YAML `lang:` → (2) `_quarto.yml` `lang:` → (3) `getOption("qpost.lang")` → omitted |
| **license** | (1) Post YAML `license:` → (2) `getOption("qpost.license")` → omitted |
| **site/blog name** | `_quarto.yml` `title:` |
| **site URL** | `_quarto.yml` `site-url:` |

This means:
- If your `_quarto.yml` already specifies a language, you don't need to configure `qpost.lang`
- You only need to set `qpost.license` once in your `.Rprofile` to apply it to all posts
- Post-level YAML always overrides defaults

### Example .Rprofile

```r
# Project-level .Rprofile (in your blog's root directory)

# These settings apply to add_coins() for this project
options(
  qpost.lang = "en",           # Fallback language for posts without lang: in YAML
  qpost.license = "CC-BY-4.0"  # License for blog post content
)
```

---

## Editing COinS After Creation

The COinS span is just an R code chunk at the end of your post. You can:

1. **Edit it manually** — change metadata values directly in the `title` attribute
2. **Regenerate it** — delete the chunk and run `add_coins()` again
3. **Remove it** — delete the chunk entirely (though this defeats the purpose!)

To re-run `add_coins()` on an existing post:
1. Delete the `coins` code chunk at the bottom
2. Run `add_coins(file_path = "posts/your-post/index.qmd")`
3. Review and commit the changes

---

## Backup Mode

By default, `add_coins(backup = TRUE)` creates a backup of your post before modifying it:

```r
add_coins(file_path = "posts/my-post/index.qmd", backup = TRUE)
```

This creates a file like `posts/my-post/index.qmd.bak` as a safety measure. If something goes wrong, you can restore the original.

To disable backups:

```r
add_coins(file_path = "posts/my-post/index.qmd", backup = FALSE)
```

---

## Testing Your COinS

Once you've added COinS to a post, you can test it:

1. **Render your blog** with `quarto preview` or `quarto render`
2. **Open the post in your browser**
3. **Install Zotero** browser extension (if you haven't already) from [zotero.org](https://www.zotero.org/download/)
4. **Click the Zotero save icon** in your browser address bar
5. **Verify the metadata** — Zotero should pre-fill the title, author, date, and other fields automatically

If the save icon doesn't appear, check that:
- The HTML output contains the COinS `<span>` (view page source in your browser)
- The metadata in the `title` attribute is properly URL-encoded
- Your Zotero browser extension is enabled

---

## Integration with Academic Workflow

COinS is particularly valuable if you:

- **Write papers that cite your own blog posts** — Zotero can cite them with proper formatting
- **Build a personal knowledge base** — Zotero can organize your posts alongside other research
- **Share research** — Others can easily add your posts to their Zotero libraries
- **Track citations** — Zotero note-taking features work with your posts as regular bibliographic items

---

## Further Reading

- [OpenURL Standard (Z39.88-2004)](https://www.niso.org/standards/standard_detail.cfm?std_id=783) — The official specification
- [COinS Specification (Wayback Archive)](https://web.archive.org/web/20161223121044/http://www.ocoins.info/) — Original COinS documentation
- [Zotero Browser Extension](https://www.zotero.org/download/) — Download and documentation
- [Zotero Support: Saving Items](https://www.zotero.org/support/adding_items_to_zotero) — How Zotero captures web content

---

## See Also

- [`qpost()`](../reference/qpost.html) — Create new Quarto blog posts interactively
- [Quarto Blog Documentation](https://quarto.org/docs/websites/website-blog.html)
